\begin{declaration}
  
  I hereby confirm this work is my own, except where other works are referenced. This work has not previously been submitted to any institute, including this one.
  
  \ChapterRef{chap:Introduction} gives an overview of the content in this thesis. \ChapterRef{chap:Neutrino Physics} and \ChapterRef{chap:SBN Program} give a review of neutrino theory and details of neutrino detection using \glspl{lartpc} with a focus on the key features of the \gls{sbn} program. These chapters cover work done by many members of the wider neutrino community and have been referenced where appropriate. 
  
  \ChapterRef{chap:Energy_Reco} builds upon the reconstruction work done by a number of people as part of the \gls{larsoft} framework. In particular, the \textit{Shower Num Electrons Energy tool} and the \textit{Shower ESTAR Energy tool} have been newly developed and validated by myself.
  
  \ChapterRef{chap:osc_inputs} discusses the necessary inputs and analysis choices required for a sterile neutrino oscillation analysis. Again, many people worked on different aspects of the work presented here. Personally, I produced the most recent \nue \gls{mc} event sample and performed the study investigating the impact of the baseline parametrisation. \ChapterRef{chap:VALOR} goes on to use the inputs from \ChapterRef{chap:osc_inputs} to perform an oscillation analysis using the VALOR fitting framework. The content shown is essentially all my own and builds upon the work done by current and previous members of the VALOR group. 
  
  \begin{comment}
  Chapters 2 - Historical neutrino overview, obviously not my own work. \\
  Chapter 3 - LArTPC physics and SBN program overview. Small contribution to work on the PDS system, otherwise not my work. \\
  Chapter 4 - Many people work on Larsoft / Reconstruction. Shower Linear Energy tool originally worked on by Dom. Other two tools developed by me. \\
  Chapter 5 - \nue event production and selection is work done by Dom (and Gray?), however, the most recent events sample were produced by me using their work. GENIE/miniboone systematics is work done by SOMEONE?. Detector systematics developed by VALOR group. Baseline approximation done by steve/dom. Impact of baseline studies done by me. \\
  Chapter 6 - VALOR intially develped by Costas et. al. \nue stuff all done by me. Rhiannon did \numu. Impact of detector systematics all done by moi. 
  
  %\begin{flushright}
  %  Name??
  %\end{flushright}
  \end{comment}
\end{declaration}