%% Title
%\DTMlangsetup[en-US]{showyear=false}
\titlepage[\phantom{This text will be invisible} \\ 
\monthyeardate]
{% 
Thesis submitted in accordance with the requirements of the University of Liverpool for the degree of Doctor in Philosophy}



%% Acknowledgements
\begin{acknowledgements}
\begin{comment}
  Acknowledgements\dots \\
  Costas \\
  VALOR group as a whole \\
  Dom \& Ornella for the reco work
\end{comment}
\end{acknowledgements}

%% Abstract
\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
  %\thispagestyle{empty}
  Abstract..
\end{abstract}


%% Declaration
\begin{declaration}
  Declaration
  \vspace*{1cm}
  \begin{comment}
  Chapters 2 - Historical neutrino overview, obviously not my own work. \\
  Chapter 3 - LArTPC physics and SBN program overview. Small contribution to work on the PDS system, otherwise not my work. \\
  Chapter 4 - Many people work on Larsoft / Reconstruction. Shower Linear Energy tool originally worked on by Dom. Other two tools developed by me. \\
  Chapter 5 - \nue event production and selection is work done by Dom (and Gray?), however, the most recent events sample were produced by me using their work. GENIE/miniboone systematics is work done by SOMEONE?. Detector systematics developed by VALOR group. Baseline approximation done by steve/dom. Impact of baseline studies done by me. \\
  Chapter 6 - VALOR intially develped by Costas et. al. \nue stuff all done by me. Rhiannon did \numu. Impact of detector systematics all done by moi. 
  
  %\begin{flushright}
  %  Name??
  %\end{flushright}
  \end{comment}
\end{declaration}





%% Preface
%\begin{preface}
%  Preface??
%\end{preface}

%% ToC
\tableofcontents


%% Strictly optional!
%\frontquote{%
%  Writing in English is the most ingenious torture\\
%  ever devised for sins committed in previous lives.}%
%  {James Joyce}
%% I don't want a page number on the following blank page either.
\thispagestyle{empty}

%% Add Glossary
\glsaddall 
\printglossaries

% Have the short form for glossary entries in the list of tables and figures
\glsunsetall

%% I prefer to put these tables here rather than making the
%% front matter seemingly interminable. No-one cares, anyway!
\listoffigures
\listoftables

% Reset the glossary so the definitions start after list of tables and figures
\glsresetall



